---
title: "data-viz"
author: "Peter Menzies"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# general use
library(tidyverse)
library(here)
library(janitor)

# dataviz
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggbeeswarm)
library(ggExtra)

library(gapminder)

library(sf)
```

## Get the lizards data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

## Make the worlds worst histogram of lizard weights

```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "light blue",
                 color = "purple",
                 size = 0.5,
                 linetype = "dotted")
```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 2,
             fill = "black",
             alpha = 1,
             color = "pink",
             size = 4) +
  theme_bw()
```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, size = total_length),
             shape = 22,
             fill = "black")
```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name)
```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail)
```

## Getting things in order

Find total lizard counts by common_name

```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(n = n())

# another way to do this (awesome if you JUST want to find counts, but the previous method is necessary if you want to do more operations than just count)

lizard_counts <- lizards %>% count(common_name)
```

Convert common_name to an ordered factor

```{r}
lizard_counts <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, n))
```


```{r}
ggplot(lizard_counts, aes(y = fct_reorder(common_name, n), x = n)) +
  geom_col(aes(fill = common_name)) +
  labs(x = "Number of individuals", 
       y = "Common name", 
       title = "Lizard Counts")
```

## Axis scale breaks, limits, and labels

Scatterplot of total length vs weight

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 450, 50),
                     limits = c(0, 460),
                     expand = c(0, 5)) +
  scale_y_continuous(expand = c(0, 2))

```
Transform the date column to class Date, then find counts of observations by date

```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  group_by(date) %>% 
  summarize(count = n())
```

```{r}
ggplot(lizard_counts, aes(x = date, y = count)) +
  geom_line(color = "blue") +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y")
```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("green", "blue", "purple", "pink"), 
                     breaks = c(0, 1, 2, 3, 4, 5, 10, 20, 40 , 50))
```

### Update a color scheme using a pallette in palletteer

```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, median))

ggplot(lizards_fct, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "beyonce::X125")
```

#### Within `theme()`, the thing with the highest resolution will take precedence (e.g. `panel.grid.major.x` carries more weight than `panel.grid.major`)

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.grid.major.x = element_line(color = "red"),
    panel.grid.major = element_line(color = "blue"),
    plot.background = element_rect(fill = "cyan4"),
    axis.text.x = element_text(color = "pink"),
    panel.background = element_blank()
  )
```

## ggrepel

Make a subset from lizards that only contains obs for western whiptail lizards from the site "sand"

```{r}
ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail", site == "sand")

ggplot(ww_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num), size = 3)
```

Create a scatterplot of gdp per capita and life exp labeled by country name

```{r}
europe_gap <- gapminder %>% 
  filter(continent == "Europe", year == 2002)

ggplot(europe_gap, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country))
```

## gghighlight

Specifies conditions where points or series will be highlighted from the others - also automatically grays out the non-highlighted values

```{r}
p <- ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point()

p + gghighlight(toe_num == 250, label_key = toe_num)
```

```{r}
q <- ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name)) +
  gghighlight(max(weight) > 30, label_key = common_name)

q
```

## Patchword for compound figures

```{r}
# follows order of op - and you can apply theme to all with &

(p | q) / q &
  theme_minimal()
```

## A few new graph types

### Marginal plots

```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight)

ggplot(whiptails, aes(x = total_length, weight)) +
  geom_point() +
  geom_rug()


```

```{r}
my_plot <- ggplot(whiptails, aes(x = total_length, weight)) +
  geom_point(aes(color = sex), size = 2) +
  scale_color_manual(values = c("cyan4", "black", "goldenrod"),
                     name = "Sex: ",
                     labels = c("female", "juvenile", "male")) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggMarginal(my_plot, type = "boxplot", groupColour = TRUE)
```
### Beeswarm plot

```{r}
ggplot(whiptails, aes(x = sex, y = weight)) +
  geom_beeswarm() +
  geom_boxplot(fill = NA)
```

### Heatmap with geom_tile()

```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  mutate(year = lubridate::year(date)) %>% 
  count(year, common_name)
```


```{r}
# heatmap with geom_tile

ggplot(lizard_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = n), color = "white", size = 3) +
  scale_fill_gradientn(colors = c("navy", "red", "orange")) +
  # or - colorblind-friendly built in theme (c for continuous)
  scale_fill_viridis_c()

```

### make a map!

```{r}
# reading in a kml file - contains spatial information

vegetation <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(Name) %>% 
  clean_names()

ggplot(vegetation) +
  geom_sf(aes(fill = name), color = "white", size = 0.2) +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
  labs(x = "Longitude", y = "Latitude", fill = "Dominant vegetation:") +
  theme_void()
```



